---
title: "Homework 2"
author: "Jing Leng"
date: "September 24, 2014"
output: pdf_document
---
```
# STATS 406 HW2 
setwd("~/Documents/GoBlue/F14/F14R")
mix <- read.table("mixgauss.dat")
head(mix)
dim(mix)
library(rgl)

# 1 
## a 
plot3d(mix, type = 'p', size = 10) 
## b
pairs(mix) 
## c
c1 <- c(-2, -2, -4) # observed centers
c2 <- c(-4, -2, 2)
c3 <- c(4, 6, 2)

centers <- rbind(c1, c2, c3)

classify <- function(x) {
  dis <- apply(centers, 1, function(y) sqrt(sum((x - y)^2))); 
  which(dis == min(dis))
}
label <- apply(mix, 1, classify )

mix$class <- label
plot3d(mix, type = 'p', col = mix$class + 1, size = 10) 
plot3d(centers, type = 'p', col = 'blue', size = 20, add = T)
pairs(mix[, 1:3], col = mix$class + 1, pch = ifelse(mix$class < 3, mix$class -1, mix$class + 2))

p1 = sum(mix$class == 1)/dim(mix)[1]
p2 = sum(mix$class == 2)/dim(mix)[1]
p3 = 1 - p1 -p2
p1; p2; p3

# 2
## a
lai <- read.csv('LAI.csv', header = F)
agg <- apply(lai, 1, sum)
which(agg == max(agg))

maxmonths <- sapply(0:5, function(i) which.max(agg[(12*i+1):(12*i+12)])) #use sapply to get the months with largest LAI.
maxmonths 
minmonths <- sapply(0:5, function(i) which.min(agg[(12*i+1):(12*i+12)])) #use sapply to get the months with largest LAI.
minmonths 

yearcol = matrix(agg,nrow=12)
#now with the above, using apply(), how do you find the yearly aggregated LAI over the continent?
 yragg = apply(yearcol, 2, sum) #along which axis?(2-column), by which function?(sum)

maxyear <- which(yragg == max(yragg))
minyear <- which(yragg == min(yragg))

lai<-as.matrix(lai)
maxt<-dim(lai)[1] # 72 time points
LAI<-sapply(1:maxt,function(i) matrix(lai[i,],nrow=120,ncol=60,byrow=T),simplify="array")

## b 
LAImean <- apply(LAI, 1:2, mean)
LAIsd <- apply(LAI, 1:2, sd)
dim(LAI)
dim(LAImean)


## c
image.plot(LAImean, horizontal = T)
image.plot(LAIsd, horizontal = T)

image.plot(LAI[80:90,25:45,7],horizontal=T) ## used to determine a location within Michigan. 
image.plot(LAI[48:80,10:30,7],horizontal=T) ## used to determine a location within Michigan. 
image.plot(LAI[,,7],horizontal=T) ## used to determine a location within Michigan. 

# Michigan [87, 33]
# Arizona [57,28]
plot(LAI[87, 33,], type = 'l', ylim = c(0, 4), xlab = 'month', ylab = 'lai')
par(new = T)
plot(LAI[57, 28,], type = 'l', ylim = c(0, 4), xlab = '', ylab = '', lty = 2)
## d
maxloc <- which(LAImean == max(LAImean))
minloc <- which(LAImean == min(LAImean[LAImean != 0]))
maxloc <- c(maxloc%/%120, maxloc%%120)
minloc <- c(minloc%/%120, minloc%%120)
maxloc; minloc; 

## e
leftlim <- (2005-2000)*12+1
rightlim <- (2005-2000)*12+12
par(ask = T)
for (i in leftlim:rightlim) {
  image.plot(LAI[65:100,15:35,i],horizontal=T) 
}

```
##1
###a).    
  
```{r echo = F, fig.height=4, fig.width=5}
setwd("~/Documents/GoBlue/F14/F14R")
mix <- read.table("mixgauss.dat")
library(scatterplot3d)
scatterplot3d(mix)
```
  
Yes, there are three clusters of data points.   
  
###b).   
```{r echo = F, fig.height=4, fig.width=5}
pairs(mix)
```
The coordinates for the three cluster centers are estimated to be: (-2, -2, -4), (-4, -2,
2) and (4, 6, 2)

###c). 
####3-d
```{r echo = F, fig.height=4, fig.width=5}
library(scatterplot3d)
c1 <- c(-2, -2, -4) # observed centers
c2 <- c(-4, -2, 2)
c3 <- c(4, 6, 2)

centers <- rbind(c1, c2, c3)

classify <- function(x) {
  dis <- apply(centers, 1, function(y) sqrt(sum((x - y)^2))); 
  which(dis == min(dis))
}
label <- apply(mix, 1, classify )

mix$class <- label
scatterplot3d(mix[,1:3], color = mix$class + 1, pch = ifelse(mix$class < 3, mix$class -1, mix$class + 2))
``` 

####2-d
```{r echo = F, fig.height=4, fig.width=5}
pairs(mix[, 1:3], col = mix$class + 1, pch = ifelse(mix$class < 3, mix$class -1, mix$class + 2))

p1 = sum(mix$class == 1)/dim(mix)[1]
p2 = sum(mix$class == 2)/dim(mix)[1]
p3 = 1 - p1 -p2
```
  
The proportion of the three clusters are: `r p1`, `r p2`, `r p3`. 

##2 
###a) 
```{r echo = FALSE}
lai <- read.csv('LAI.csv', header = F)
agg <- apply(lai, 1, sum)

maxmonths <- sapply(0:5, function(i) which.max(agg[(12*i+1):(12*i+12)])) #use sapply to get the months with largest LAI.
minmonths <- sapply(0:5, function(i) which.min(agg[(12*i+1):(12*i+12)])) #use sapply to get the months with largest LAI.

yearcol = matrix(agg,nrow=12)
#now with the above, using apply(), how do you find the yearly aggregated LAI over the continent?
 yragg = apply(yearcol, 2, sum) #along which axis?(2-column), by which function?(sum)
maxyear <- which(yragg == max(yragg))
minyear <- which(yragg == min(yragg))

lai<-as.matrix(lai)
maxt<-dim(lai)[1] # 72 time points
LAI<-sapply(1:maxt,function(i) matrix(lai[i,],nrow=120,ncol=60,byrow=T),simplify="array")

## b 
LAImean <- apply(LAI, 1:2, mean)
LAIsd <- apply(LAI, 1:2, sd)
```

For each year from 2000 to 2005, the month with maximum LAI is (respectively): `r maxmonths`.  The month with minimum LAI is (respectively): `r minmonths`. 
The year with maximun aggregated LAI is: `r maxyear+2000`, with minimum: `r minyear+2000`. 

###b)  
####mean 
```{r echo = F }
library(fields)
image.plot(LAImean, horizontal = T)
```

####sd 
```{r echo = F}
image.plot(LAIsd, horizontal = T)
``` 

###c) 

```{r echo =FALSE}
plot(LAI[87, 33,], type = 'l', ylim = c(0, 4), xlab = 'month', ylab = 'lai')
par(new = T)
plot(LAI[57, 28,], type = 'l', ylim = c(0, 4), xlab = '', ylab = '', lty = 2)
```
  
SE Michigan (location: 87, 33) has higher LAI than Arizona (location: 57, 28). 

###d) 
```{r echo =FALSE}
maxloc <- which(LAImean == max(LAImean))
minloc <- which(LAImean == min(LAImean[LAImean != 0]))
maxloc <- c(maxloc%/%120, maxloc%%120)
minloc <- c(minloc%/%120, minloc%%120)
``` 
Most leafy: `r maxloc`.  
Least leafy: `r minloc`. 

###e) 
```{r echo = F,fig.width = 4, fig.height=3.7} 
leftlim <- (2005-2000)*12+1
rightlim <- (2005-2000)*12+12
par(ask = T)
for (i in leftlim:rightlim) {
  image.plot(LAI[65:100,15:35,i],horizontal=T) 
}
```