---
title: "Homework 3"
author: "Jing Leng"
date: "September 25, 2014"
output: pdf_document
---
##1 
###a)
```{r}
library(MASS)
mu <- c(3,4)
sigma <- matrix(c(1.0,0.8,0.8,1.0),nrow=2)
datam <- data.frame(mvrnorm(100,mu,sigma))
colnames(datam) <- c("x","y")

plot(y~x, datam, xlim=c(0,7),ylim=c(0,7))
lm <- lm(y~x, datam)
abline(lm)
lm2 <- lm(x~y, datam)
lines(lm2$fitted.values, datam$y)
```

###b) 
```{r}

summary(lm)
summary(lm2)
```


The $R^{2}$, t-statistcs, F-statistcs and their respective p-values are the same.  

###c)
$$
\begin{aligned}
T_{y|x}&=\frac{\sqrt{S_{xx}}(\hat{\beta}-\beta)}{\hat{\sigma}}
=\frac{\sqrt{S_{xx}}(\frac{S_{xy}}{S_{xx}}-\beta)}{\sqrt{S_{yy}-\frac{S_{xy}^{2}}{X_{xx}}}}
=\frac{S_{xy}-S_{xx}\beta}{\sqrt{S_{yy}S_{xx}-S_{xy}^{2}}}
=\frac{S_{xy}}{\sqrt{S_{yy}S_{xx}-S_{xy}^{2}}} 
\end{aligned} 
$$
For $T_{y|x}$, x and y are interchangeable, thus $T_{x|y}=T_{y|x}$.  

###d). 
  
  
  
  
  
  
  
  
###e)
```{r}
left <- (lm$fitted.values - mean(datam$y))/(sd(datam$y))
r = sqrt(1-(sum(lm$residuals^2))/(sum((datam$y - mean(datam$y))^2)))
right <- r*(datam$x - mean(datam$x))/(sd(datam$x))
sum(abs(left - right))
```
First 5 elements for left: `r left[1:5]`.  
First 5 elements for right: `r right[1:5]`. 

##2 
###a)
```{r}
college <- read.csv('college.csv')
head(college)
pairs(college[c(1,2,4,5,6)])
```

###b) 
```{r}
lm2 <- lm(gradrat ~ . - lenroll - private - stufac - rmbrd, data = college) 
anova(lm, lm2)
``` 
  
F-statistics is 1.1251, p-value is 0.342. We fail to reject the null hypothesis. 
###c)
```{r}
summary(lm)$coefficients
coe <- summary(lm)$coefficients
tstats <- (coe[2, 1] - 0.05)/coe[2,2]
tstats
pv <- 2*pt((coe[2, 1] - 0.05)/coe[2,2], 120 - 6)
pv
```
  
t-statistics is `r tstats`, p-value is `r pv`. We fail to reject the null hypothesis. 

###d)
```{r}
lm3 <- lm(gradrat ~ . - lenroll - private - act, data = college)
anova(lm, lm3)
```
The F-statistics is 45.289, p-value is close to 0. We reject the null hypothesis. 

###e)
```{r}
lm4 <- lm(gradrat ~ I(rmbrd + act) + csat + private + stufac, college)
anova(lm, lm4)
```
The F-statistics is 7.8216, p-value is 0.0061. We reject the null hypothesis. 

##3
```{r}
x <- c(1, 0, -1)
y <- c(1, 0, 2)
lm <- lm(y ~ x + I(3*x^2 -2))
lm2 <- lm(y ~ x)
summary(lm)
summary(lm2)
```
$\beta_{0}$ and $\beta_{1}$ remain unchanged no matter whether $\beta_{2}=0$. 