---
title: "Homework 4"
author: "Jing Leng"
date: "October 2, 2014"
output: pdf_document
---
##1 
```{r echo = F}
library(faraway)
```
###a) 
```{r}
lm <- lm(lpsa~., prostate)
summary(lm)$coefficient
confint(lm)["age",]
confint(lm, level = 0.9)["age",]
```
0 is in the 95% confidence interval of parameter for `age`, we fail to reject the null hypothesis. The p-value for variable `age` is 0.08229 > 0.05, therefore we fail to reject the null hypothesis.  
0 is not in the 90% confidence interval of parameter for `age`, we reject the null hypothesis. The p-value 0.08229 < 0.1, we reject the null hypothesis. 

###b) 
```{r fig.width=5, fig.height=4}
library(ellipse)
plot(ellipse(lm, c("age", "lbph")),
     type="l")
points(lm$coef["age"],lm$coef["lbph"],pch=18)
points(0,0)
```
  
The null hypothesis is: 
$$H_{0}:\beta_{age}=\beta_{lbph}=0$$
Since the origin is within the ellipse, we fail to reject the null hypothesis. 

###c) 
```{r}
newdata <- data.frame(lcavol = 1.44692, lweight = 3.62301, age = 65, lbph = 0.3001, 
                      svi = 0, lcp = -0.79851, gleason = 7, pgg45 = 15)

predict(lm, newdata, interval = "confidence")
predict(lm, newdata, interval = "prediction")
```
The confidence interval for mean response is (2.17, 2.61), the prediction interval for new response is (0.96, 3.81). 

##2 
###a) 
```{r fig.width=5, fig.height=4}
lm2 <- lm(gamble ~ ., data = teengamb)
st_res <- rstandard(lm2)
plot(fitted.values(lm2),st_res,
     xlab="Fitted values",
     ylab="Standardized residuals")
abline(h = 0)
```
The constant variance assumption does not hold. We transform the response variable into square root of `gamble`. 
```{r fig.width=5, fig.height=4}
newgamble <- sqrt(teengamb$gamble)
teengamb$newgamble <- newgamble
lm3 <- lm(newgamble ~ . - gamble, data = teengamb)

st_res <- rstandard(lm3)
plot(fitted.values(lm3),st_res,
     xlab="Fitted values",
     ylab="Standardized residuals")
abline(h = 0)
``` 
The constant variance assumption holds. We will continue with the new model.  

###b) 
```{r fig.width=5, fig.height=4}
qqnorm(lm3$residual, ylab="Residuals")
qqline(lm3$residual)
```
Residuals can be seen as normaly distributed. 

###c) 
```{r fig.width=5, fig.height=4}
halfnorm(influence(lm3)$hat, nlab = 2, ylab="Leverages")
teengamb[c(42, 35), ]
```
Number 42 and number 35 has the largest leverages. 

###d) 
```{r}
ti <- rstudent(lm3)
max(abs(ti))
which(abs(ti) == max(abs(ti)))
p <- 2*(1-pt(max(abs(ti)), df=47-4-1))
thres <- 0.05/47 
```
The p-value = `r p` > `r thres`, we fail to reject the null hypothsis. Therefore we cannot name a outlier from the data. 

###e) 
```{r fig.width=5, fig.height=4}
cook <- cooks.distance(lm3)
halfnorm(cook, nlab = 3, ylab="Cooks distance")
```
  
number 5, 39 and 24 are suspected to be influential points. 
```{r}
lm4 <- lm(newgamble ~ . - gamble, data = teengamb[-24, ])
predict(lm4, teengamb[24,], interval ="prediction")
teengamb[24,]$newgamble

lm5 <- lm(newgamble ~ . - gamble, data = teengamb[-39, ])
predict(lm5, teengamb[39,], interval ="prediction")
teengamb[39,]$newgamble


lm6 <- lm(newgamble ~ . - gamble, data = teengamb[-5, ])
predict(lm6, teengamb[5,], interval ="prediction")
teengamb[5,]$newgamble
```

The real response for 24 and 39 are not in the respective prediction intervals in models excluding them. Therefore 24 and 39 are influential points. The real response for 5 is in the prediction interval, so it is not influential. 